---
title: k-Nearest Neighbours predictions
author: Rémi NICOLE
output:
  pdf_document:
    fig_caption: yes
...

Introduction
============

```{r}
library(ggplot2)
library(knitr)
library(class)
```

```{r}
myData <- read.csv("01_heights_weights_genders.csv")
# Convert inches to centimeters
myData$Height <- sapply(myData$Height, function(x) return(x*2.54))
# Convert pounds to kilograms
myData$Weight <- sapply(myData$Weight, function(x) return(x/2.205))
```

```{r, echo=FALSE}
kable(head(myData), caption="Sample of Heights and Weights data")
kable(summary(myData), caption="Summary of Heights and Weights data")
```

Normalizing
===========

```{r}
normalize <- function(x) {
	num <- x - min(x)
	den <- max(x) - min(x)
	return(num/den)
}
```

```{r}
myNormalizedData <- as.data.frame(lapply(myData[2:3], normalize))
myNormalizedData$Gender <- myData$Gender
```

```{r, echo=FALSE}
kable(head(myNormalizedData), caption="Sample of normalized Heights and Weights data")
kable(summary(myNormalizedData), caption="Summary of normalized Heights and Weights data")
```

The training set
================

We divide the data into two shuffled groups, one will be used to train the
system, and the other one is to test the trained system.

```{r}
ind <- sample(2, nrow(myNormalizedData), replace=TRUE, prob=c(0.67, 0.33))

myNormalizedData.training <- myNormalizedData[ind==1, 1:2]
myNormalizedData.trainLabels <- myNormalizedData[ind==1, 3]

myNormalizedData.test <- myNormalizedData[ind==2, 1:2]
myNormalizedData.testLabels <- myNormalizedData[ind==2, 3]
```

```{r, echo=FALSE}
kable(head(myNormalizedData.training), caption="Sample of normalized Heights and Weights from the training data")
kable(head(myNormalizedData.test), caption="Sample of normalized Heights and Weights from the test data")
```

KNN
===

```{r}
predictions <- knn(train = myNormalizedData.training,
                   test  = myNormalizedData.test,
                   cl    = myNormalizedData.trainLabels,
                   k     = 3)
results <- data.frame(Predictions=predictions,
                      Actual.Results=myNormalizedData.testLabels)
incoherences <- sum(results$Predictions != results$Actual.Results)
```

```{r, echo=FALSE}
kable(head(results, n=20), caption="Sample of the comparison between $knn$'s predictions and the actuel results")
```

Here, we found a total number of incoherent prediction of
\(\Delta = `r incoherences`\), which means a error ratio of
\(\dfrac{\Delta}{n} = \dfrac{`r incoherences`}{`r length(results$Actual.Results)`} = `r round(100* incoherences / length(results$Actual.Results), digits=2)`\%\).

```{r, fig.cap="k-nearest neighbours visual comparison"}
femaleTruePredictions <- ((results$Predictions    == "Female") &
                          (results$Actual.Results == "Female"))*1

femaleWrongPredictions <- ((results$Predictions    == "Female") &
                           (results$Actual.Results == "Male"))*2

maleTruePredictions <- ((results$Predictions    == "Male") &
                        (results$Actual.Results == "Male"))*3

maleWrongPredictions <- ((results$Predictions    == "Male") &
                         (results$Actual.Results == "Female"))*4

predictionsCategories <- factor(
	femaleTruePredictions +
	femaleWrongPredictions +
	maleTruePredictions +
	maleWrongPredictions)

levels(predictionsCategories) <- c(
	"Female true prediction",
	"Female wrong prediction",
	"Male true prediction",
	"Male wrong prediction")

ggplot(myNormalizedData.test, aes(x=Height, y=Weight)) +
	geom_point(size=.8, aes(color=predictionsCategories)) +
	scale_color_discrete(name="Predictions") +
	theme(legend.position=c(0.83,0.25))
```
