---
title: Text Analysis and Data Clustering
author: Rémi NICOLE
output:
  pdf_document:
    fig_caption: yes
...

Introduction
============

```{r "libraries"}
library(tm)
data(acq)
library(SnowballC)
library(knitr)
library(ggplot2)
library(gplots)
library(RColorBrewer)
library(wordcloud)
library(cluster)
library(fpc)
library(topicmodels)
```

Cleaning and Stemming
=====================

In order to clean the text, we have to remove the punctuation, the numbers,
the stop-words of the English language and convert the text into all-lowercase
text.

We will also "stem" the text, which means that every word will be stripped
down to the radical of the word. For example, the word "saying" will be
replaced by "say".

```{r "clean stem"}
acq.clean <- tm_map(acq, removePunctuation)
acq.clean <- tm_map(acq.clean, removeNumbers)
acq.clean <- tm_map(acq.clean, tolower)
acq.clean <- tm_map(acq.clean, removeWords, c(stopwords("english"), "said"))
acq.clean <- tm_map(acq.clean, stemDocument)
acq.clean <- tm_map(acq.clean, stripWhitespace)
acq.clean <- tm_map(acq.clean, PlainTextDocument)
# Print a sample of the cleaned text
str(acq.clean[1][[1]]$content, width=80, strict.width="cut")
```

Document term matrices
======================

Then, we transform that data into a `DocumentTermMatrix` which is a matrix of
occurrences of all the words stored in the data. We then sum the columns of
the matrix to obtain the number of occurrences of the words.

```{r "dtm"}
acq.dtm <- DocumentTermMatrix(acq.clean)
acq.freqs <- colSums(as.matrix(acq.dtm))
```

```{r "top frequencies", echo=FALSE}
kable(head(acq.freqs[order(acq.freqs, decreasing=TRUE)], n=10), caption="Top frequent words in the acq dataset")
```

```{r "histogram", fig.cap="Histogram of the most used words"}
acq.freqs.small <- acq.freqs[tail(order(acq.freqs), n=25)]
acq.freqs.small <- data.frame(Word=names(acq.freqs.small),
                              Frequency=acq.freqs.small)
ggplot(acq.freqs.small, aes(Word, Frequency)) +
	geom_bar(stat="identity") +
	theme(axis.text.x=element_text(angle=45, hjust=1))
```

```{r "wordcloud", fig.cap="Wordcloud of the acq dataset"}
acq.freqs <- data.frame(Word=names(acq.freqs),
                        Frequency=acq.freqs)
wordcloud(acq.freqs$Word,
          acq.freqs$Frequency,
          max.words=100,
          scale=c(5, .1),
          colors=brewer.pal(9, "Blues")[4:9])
```

Hierarchical clustering
=======================

From this part on, we will remove the sparse words. As our matrix is very
sparse, we will only decrease the sparsity to 80% to have enough data to have
meaningful clustering.

```{r "distances and fitting"}
acq.dtms <- removeSparseTerms(acq.dtm, .8)
acq.distances <- dist(t(acq.dtms), method="euclidian")
acq.hfit <- hclust(acq.distances, method="ward.D2")
```

```{r "heatmap", fig.cap="Heatmap of the distance between the words"}
heatmap.2(as.matrix(acq.distances), trace="none")
```

```{r "dendogram", fig.cap="Clustered dendogram of the relations between the words"}
plot(acq.hfit, hang=-1)
rect.hclust(acq.hfit, k=7, border="gray")
```

K-means clustering
==================

```{r "cluster plot"}
acq.kfit <- kmeans(acq.distances, 5, 500)
clusplot(as.matrix(acq.distances), acq.kfit$cluster, color=T, shade=T, labels=2)
```

Topic models
============

```{r "topic finding"}
acq.lda <- LDA(acq.dtm, k=7)
acq.terms <- terms(acq.lda, 4)
acq.terms <- apply(acq.terms, MARGIN=2, paste, collapse=", ")

acq.topics <- data.frame(document=1:50, topics(acq.lda))
```

```{r "Topic terms table", echo=FALSE}
kable(acq.terms, caption="Topic terms")
```

```{r "document topics"}
acq.topicsWKeywods <-
	data.frame(Document=1:50,
	           Topic.Keywords=acq.terms[acq.topics$topics.acq.lda.])
kable(acq.topicsWKeywods, caption="Topic terms")
```

<!-- Thank you: https://rstudio-pubs-static.s3.amazonaws.com/31867_8236987cf0a8444e962ccd2aec46d9c3.html -->
<!-- Thank you: http://fr.slideshare.net/rdatamining/text-mining-with-r-an-analysis-of-twitter-data -->
